---
title: "Shiny with R Markdown"
author: ""
date: "4-15-2021"
output: html_document
editor_options: 
  chunk_output_type: console
runtime: shiny
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
  message = FALSE,
  warning = FALSE
)
```

### Main ideas

### Coming up

- Complete course evaluation in DukeHub
  - If this gets above 50%, I will add 0.125 to everyone's final grade, if we get to 65%, I will add 0.25.
- Project work
  - I will send out information about a **very short** peer evaluation survey later in the week. 

### Shiny with R Markdown

```{r load-packages}
library(tidyverse)
library(shiny)
```

So far we have used R Markdown to create static reports for homeworks, labs, 
and lecture notes. Today we are going to make R Markdown documents interactive.
Readers will be able to interact with your report and see the results 
immediately. You will also be able to share reports with friends and family 
without them having to download `R` or `RStudio`.

Note that the YAML header at the top of this file contains `runtime: shiny` and
instead of the "Knit" button there is a "Run Document" button.

Below are a few examples of Shiny applets. Choose one or two to run and try
interacting with the applet. Then, examine `app.R` on the bottom of the page.

**Question:** What do you notice when comparing the applet and `app.R`?

```{r examples-shiny, eval = FALSE}
runExample("01_hello") # a histogram
runExample("02_text") # tables and data frames
runExample("03_reactivity") # a reactive expression
runExample("04_mpg") # global variables
runExample("05_sliders") # slider bars
runExample("06_tabsets") # tabbed panels
runExample("07_widgets") # help text and submit buttons
runExample("08_html") # Shiny app built from HTML
runExample("09_upload") # file upload wizard
runExample("10_download") # file download wizard
runExample("11_timer") # an automated timer
```

Now try clicking "Run Document" at the top of this page.

### Inputs and Outputs

Shiny **inputs** are values obtained from the user. Shiny **outputs** are 
updated automatically whenever inputs change. 

In the code chunk, below we make a basic R plot interactive by wrapping it in 
the Shiny `renderPlot()` function.

The functions `selectInput()` and `sliderInput()` create the widgets that a user
employs to adjust the plot.

`selectInput()` has the name `n_bins` and allows the user to specify the number
of bins in the histogram.

`sliderInput()` has the name `alpha_adjust` and allows the user to specify the
transparency.

When generating the plot, we refer to these inputs as `input$n_bins` and 
`input$alpha_adjust`.

```{r shiny-plot-1, echo=FALSE}
inputPanel(
  selectInput("n_bins",
    label = "Number of bins:",
    choices = c(5, 10, 15, 20, 50), selected = 20
  ),

  sliderInput("alpha_adjust",
    label = "Alpha adjustment:",
    min = 0, max = 1, value = 1, step = .1
  )
)

renderPlot({
  ggplot(mpg, aes(x = cty)) +
    geom_histogram(
      bins = input$n_bins,
      alpha = input$alpha_adjust
    ) +
    labs(
      x = "City Miles per Gallon",
      y = "",
      title = "Histogram of Car Mileage"
    )
})
```

Using the template below, we  to create a Shiny app that generates a 
scatterplot of `cty` versus `hwy` that allows users to color the points by 
class.

```{r shiny-plot-2, echo=FALSE, warning=FALSE}
base_ggplot <- ggplot(mpg, aes(x = hwy, y = cty))

inputPanel(
  checkboxInput("type_check", "Color by car type?", value = FALSE)
)
renderPlot({
  if (input$type_check) {
    base_ggplot +
      geom_jitter(aes(color = class)) +
      labs(
        x = "Highway miles per gallon",
        y = "City miles per gallon",
        title = "City versus highway mileage"
      ) +
      theme_bw()
  } else if (!input$type_check) {
    base_ggplot + geom_jitter() +
      labs(
        x = "Highway miles per gallon",
        y = "City miles per gallon",
        title = "City versus highway mileage"
      ) +
      theme_bw()
  }
})
```

Helpful inputs for collecting values from the user are provided below. See the
Shiny cheatsheet and additional resources for more information.

- `checkboxGroupInput()`
- `checkboxInput()`
- `fileInput()`
- `numericInput()`
- `radioButtons()`
- `selectInput()`
- `sliderInput()`
- `textInput()`

### Sharing Shiny Apps

What if you wanted to share a Shiny applet without having a user download and 
install `R` and `RStudio`?

**Method #1**

1. Create your Shiny app in an RMarkdown file. A template document
  `first_shiny.Rmd` has already been created for you.
  
2. Create a new folder for your Shiny app (shiny_app)

3. Create a new `R` script called `app.R` in your shiny_app folder. If you are 
  using your own data, include it in a separate `data` folder in your app 
  folder.

4. Cut and paste the code from `first_shiny.Rmd` into `app.R` and convert your
  code to ui-server format. You will have to add some ui wrapper code to your
  inputs and some server wrapper code to your outputs. Check out the example 
  code and the additional resources for help here.
  
Now you're ready to publish!

5. Load the `rsconnect` package. 

```{r load-rsconnect, eval = FALSE}
library(rsconnect)
```

6. Authorize your account. Go to the website
  [here](https://www.shinyapps.io/admin/) and log in with your GitHub password.
  Then authorize your computer by copying and pasting the `setAccountInfo` code
  into the console. You can access this by clicking "Tokens" in the upper-right
  drop-down menu.
  
```{r run-app, eval = FALSE}
runApp("shiny_app", display.mode = "showcase")
```

7. Type the code below to deploy your app. If you are using the Docker 
  containers this will take several minutes. I recommend testing by clicking
  "Run Document" and only deploying once when you are done.

```{r deploy-app, eval = FALSE}
deployApp("shiny_app")
```

After a few minutes, a browser window will open with your new application. 

**Method #2**

1. Create an R Markdown file to produce a shiny app and use `runtime: shiny` and
`output: html_document`.

2. Save any data to the same folder as the R Markdown file and read in the data.

3. Hit Run Document.

4. Hit the Publish icon in the upper right of the HTML viewer window and select
  Publish.
  
5. Under Publish from Account, choose your shinyapps.io account. Make sure to 
  set your token information.
  
6. Check your R Markdown and data files.

7. Hit publish and obtain your URL to share.

### Practice- this is optional, just if you want to practice using Shiny, I know you have a lot of work this week!

Build on `first_shiny.Rmd` and `app.R` to create an effective applet that takes
user inputs and visualizes coronavirus cases at Duke. You can consider the 
elements below but be creative!

- an input for selecting tests or positives
- an input for selecting weekly or cumulative cases
- an input for selecting semester 1 or semester 2 data
- an input for selecting the positivity rate
- output that displays the raw data
- output that displays multiple plots
- output that displays multiple panes

### Additional Resources

- [Shiny Cheatsheet](https://shiny.rstudio.com/images/shiny-cheatsheet.pdf)
- [Authoring Shiny](https://rmarkdown.rstudio.com/authoring_shiny.html)
- [Shiny Tutorials](https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/)
- [Datacamp on R Shiny](https://www.datacamp.com/courses/building-web-applications-in-r-with-shiny)
- [Shiny Website](http://www.shinyapps.io/)
